
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# A4 - Visualisation

## Instructions
This is a skeleton markdown file that you will use to complete Assignment 3 this semester. Use the chunks (see below) to code in. Where you see ........................ please remove and write your response to the prompt. 
```{r}
###################### THIS IS A CHUNK #############################
```

As usual, follow the steps below and submit your markdown document as a pdf. You will need to knit this to a pdf or knit it to an html then save as a pdf to submit. 

YOUR MUST NAME YOUR FILE LIKE THIS: course_section_semester_assignment_unityID
For example: DSA295_003_SP25_A2_trleppar

Grading: Full marks are awarded for completion and proper naming. One mark will be dropped if not named properly, and/or additional marks dropped for each portion incomplete. This assignment will have 5 points per portion (4 prompts to complete) and a total of 20 marks. 

## Place your library() here. You will need to list each one you will use!
```{r}

```

## Bringing in your data
By this time, you're a dab hand at bringing in your network data. Please bring in your network, convert it to a network object and view that object. Just a reminder, to view the network, just list the name of the network. If my igraph object (network object) is called 'graph' I can simply code graph in the chunk and I will view it. 

```{r}

```

Clean and or transform your network data if you need/want to and provide a raw (unaltered) visualisation using the plot() function. 
```{r}

```

## Basic visualisation 
In class, we have covered multiple basic visualizations. Using the markdown files and your notes from our class about basic network visualisation, I want you to demonstrate your skill in basic network visualisation. 

Please alter the labels by either changing their size or toggling them.
```{r}

```

Change the colours of the network (nodes or edges).
```{r}

```

Present a basic visualisation, you feel appears clean. 
```{r}

```

Discuss what alterations you have done to the network and why you did them. 
..............................

## Intermediate visualisation
You have learned how to tell a story through network visualisation. These stories centre prominent and influential nodes in the network. 

Present a CLEAN network visual demonstrating popular (degree) nodes in your network. 
```{r}

```

Present a CLEAN network visual demonstrating influential (betweenness) nodes in your network. 
```{r}

```

Present them side-by-side sing the par() function. Consider presenting sub or main titles so I can tell which is which. 
```{r}

```

Briefly (1 - 2 sentences) discuss the stories that your visualisations tell. E.g. who are/are they many or few popular/influential nodes? 
..............................

## Advanced visualisation
You have learned how to create interactive, or animated network visualisations. 

Please produce one advanced network visualisation. Bare in mind that you will need to leave the visual as a widget in this markdown. 
```{r}

```

Briefly (1 - 2 sentences) discuss this visualisation. In what contexts may this visualisation be useful? 

# WELL DONE :)