```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Network Visualisation - Basic

```{r}
library(igraph)

grime_edge_list <- read.csv(file.choose(), header = TRUE) # select GRIME_EDGE_LIST_2008 _GRIME ONLY.csv

grime_08 <- graph_from_data_frame(d= grime_edge_list, directed = TRUE)
plot(grime_08)


grime_08_clean <- delete.edges(grime_08, E(grime_08)[which_loop(grime_08)])

plot(grime_08_clean)
```

## Layout

Random

```{r}
plot(grime_08_clean, layout = layout.random)
```

Grid

```{r}
plot(grime_08_clean, layout = layout.grid)
```

Small world/Circle

```{r}
plot(grime_08_clean, layout = layout.circle)
```

Layout by algorithms to demonstrate the relationships a bit better

```{r}
plot(grime_08_clean, layout = layout_with_dh(grime_08_clean))
plot(grime_08_clean, layout = layout_with_fr(grime_08_clean))
plot(grime_08_clean, layout = layout_with_kk(grime_08_clean))
```

## Changing colours of the graph

Vertex colours with verterx.color()

```{r}
plot(grime_08_clean, vertex.color = "pink")
```

Edge colour with edge.color()

```{r}
plot(grime_08_clean, edge.color = "pink")
```

## Changing the Sizes

Edge and vertex size

```{r}
plot(grime_08_clean, vertex.size = 5, edge.arrow.size = 0.5)
```

Edge thickness

```{r}
plot(grime_08_clean, vertex.size = 5, edge.width = 10)
```

## Titles

Main Title

```{r}
plot(grime_08_clean, vertex.size = 5, edge.arrow.size = 0.5, main = "Grime Collaborations 2008")
```

Sub title

```{r}
plot(grime_08_clean, vertex.size = 5, edge.arrow.size = 0.5, main = "Grime Collaborations", sub = "2008")
```

## Labels

Toggle

```{r}
plot(grime_08_clean, vertex.label = NA)
```

Size and Offset

```{r}
plot(grime_08_clean, vertex.label.cex = 0.5, vertex.label.dist = 2)
```

## General considerations for visualizations

Changing the aspect ratio by changing the margings of the plot window.

```{r}
plot(grime_08_clean, vertex.label = NA)
par(mar =c(0,0,0,0))
plot(grime_08_clean, vertex.label = NA)
```

We have reduced the margins all down to zero.

Set a seed to replicate the same layout every time

```{r}
par(mar =c(0,0,0,0))
set.seed(123)
plot(grime_08_clean, vertex.label = NA)
```

```{r}
par(mar =c(0,0,0,0))
set.seed(123)
plot(grime_08_clean, vertex.label = NA)
```

The layout is replicated - useful for recreating visuals you use.

## Plotting visuals next to each other

Vertical

```{r}
par(mfrow = c(2, 1))
par(mar =c(0,0,0,0))
set.seed(123)
plot(grime_08_clean, vertex.label = NA)
set.seed(123)
par(mar =c(0,0,0,0))
plot(grime_08_clean, vertex.label = NA, vertex.size = degree(grime_08_clean))
```

Horizontal

```{r}
par(mfrow = c(1, 2))
set.seed(123)
par(mar =c(0,0,0,0))
plot(grime_08_clean, vertex.label = NA)
set.seed(123)
par(mar =c(0,0,0,0))
plot(grime_08_clean, vertex.label = NA, vertex.size = degree(grime_08_clean))
```
